Gaia identifies VFTS682 as the most massive runaway star known to date

### 2018-05-11 VFTS682.tex

\author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$,
  S.~Justham$^{3,4}$, J.~Vink$^{5}$, F.~R.~Schneider$^6$,
  J.~Bestenlehner$^7$, A.~de~Koter$^{1}$}

\affil{$^{1}$Astronomical Institute Anton Pannekoek, University of
    Amsterdam, 1098 XH Amsterdam, The Netherlands \\
    $^{2}$ ESA, European Space Astronomy Centre, Apdo. de Correos 78, E-28691 Villanueva de la Ca\~nada, Madrid, Spain \\
    $^{3}$ {School of Astronomy \& Space Science, University of the Chinese
    Academy of Sciences, Beijing 100012, China}\\
     $^4$ {National Astronomical Observatories, Chinese Academy of
       Sciences, Beijing 100012, China}\\
     $^5$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}\\
     $^6$ {Department of Physics, University of Oxford, Keble Road,
       Oxford OX1 3RH, UK} \\
     $^7$ {Department of Physics and Astronomy, Hicks Building, Hounsfield Road, University of Sheffield, Sheffield S3 7RH, UK}
}

\thanks{M.~Renzo, \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}}
\date{}
\begin{abstract}
The formation and fates of the most massive stars remains one of the most elusive questions in stellar astrophysics. 
%
Previous spectroscopic studies identified VFTS682 as one of the most massive stars known. It resides in the field of the 30 Dor region in the Large Magellanic Cloud at a projected distance of 29\,pc of the star cluster R136. The lack of a detected offset in radial velocity derived from the spectra, lead to speculations whether such extreme stars can form in relative isolation. 
%
We use the unprecedented proper motions from the second Gaia data
release (DR2) to investigate the kinematics of this object. We show
that VFTS 682 is a bonafide runaway star moving with a relative
velocity of about $64\pm21\kms$ away from the star cluster R136.  The
inferred age, distance, and the magnitude and direction of the
velocity are consistent with a dynamical ejection scenario from the cluster. 
%
This would make VFTS 682 the most massive runaway star known to date.  It proves that star clusters are capable of ejecting their most massive members. We discuss the implications for the existence of stars with masses well above $100\Msun$ inside the cluster, the internal dynamics in dense star clusters and the final fate of this object.

>>>>

\author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$, J. Bestenlehner$^{x}$,  F.~R.~Schneider$^6$, J.~Vink$^{5}$ \SdM{et al.?} \\ \SdM{[ Alphabetical after Lennon. Possibly add further last-minute coauthors]},  }

  % S.~Justham$^{3,4}$,
%  A.~de~Koter$^{1}$ \todo{... add coauthors...} }

\affil{$^{1}$Astronomical Institute Anton Pannekoek, University of
    Amsterdam, 1098 XH Amsterdam, The Netherlands \\
    $^{2}$ ESA, European Space Astronomy Centre, Apdo. de Correos 78, E-28691 Villanueva de la Ca\~nada, Madrid, Spain \\
  %  $^{3}$ {School of Astronomy \& Space Science, University of the Chinese
   % Academy of Sciences, Beijing 100012, China}\\
    % $^4$ {National Astronomical Observatories, Chinese Academy of
     %  Sciences, Beijing 100012, China}\\
     $^5$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}\\
     $^6$ {Department of Physics, University of Oxford, Keble Road,
       Oxford OX1 3RH, UK} \\
     $^7$ {Department of Physics and Astronomy, Hicks Building, Hounsfield Road, University of Sheffield, Sheffield S3 7RH, UK}
}

%\thanks{Corresponding author:  M.~Renzo, \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}}
\date{}
\begin{abstract}
The formation and fates of the most massive stars remains one of the most elusive questions in stellar astrophysics. 
%
Previous spectroscopic studies identified VFTS682 as one of the most massive stars known. It resides in the field of the 30 Dor region in the Large Magellanic Cloud at a projected distance of 29\,pc of the star cluster R136. The lack of a detected offset in radial velocity derived from the spectra, lead to speculations whether such extreme stars can form in relative isolation. 
%
We use the unprecedented proper motions from the second Gaia data
release (DR2) to investigate the kinematics of this object. We show
that VFTS 682 is a bonafide runaway star moving with a relative
velocity of about $64\pm21\kms$ away from the star cluster R136.  The
inferred age, distance, and the magnitude and direction of the
velocity are consistent with a dynamical ejection scenario from the cluster. 
%
This would make VFTS 682 the most massive runaway star known to date.  It proves that star clusters are capable of ejecting their most massive members. We discuss the implications for the existence of stars with masses well above $100\Msun$ inside the cluster, the internal dynamics in dense star clusters and the final fate of this object.
